---
title: "Machine Learning Assignment"
author: "W. Lin"
date: "14 February 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
library(dplyr)
library(caret)
library(gbm)
```

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).


The goal of this project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. 

## Process:

### Preparation of clean data for building model.
```{r readdata}

if(!file.exists("../pml-training.csv")) {
        download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",
                      destfile = "../pml-training.csv")
}

dataorg <- read.csv("../pml-training.csv", header = TRUE, na.strings = c("","NA","#DIV/0!"))
```


Columns with too many empty values and those not so relevant (columnS 1 to 7) are deleted
```{r cleanvoid}
intraining <- dataorg[, 8:160]
intraining <- intraining[, apply(intraining, 2, function(x) all(!is.na(x)))]
```


There are still `r dim(intraining)[2]` variables in the data, which is a little too large dataset for running prediction. In this case, those highly correlated variables are excluded, because they can be represented by other correlated variables.


```{r cleancorr}
colnumber <- dim(intraining)[2]
corrDF <- cor(intraining[,1:colnumber -1])
resCorr <- findCorrelation(corrDF, 0.6)
intraining <- intraining[, -resCorr]
```

Till here, the number of predictors is reduced to `r dim(intraining)[2]`.


### Use Boosting and Ramdom Forest methods to build the fitting model
Both methods are learned in the lectures and they are suitable for classification like this case.

```{r mod}
inTrain <- createDataPartition(y=intraining$classe, p=0.7, list = FALSE)
training <- intraining[inTrain,]
testing <- intraining[-inTrain,]

set.seed(1235)
modegbm <- train(classe~., data = training, method = "gbm")

set.seed(1234)
moderf <- train(classe ~., data=training, method = "rf")
```

Result of the model:
```{r showmodel, echo=FALSE}
modegbm
moderf
```


### 3.Comparing two models using Confusion Matrix
```{r }
predgbm = predict(modegbm, testing)
confgbm <- confusionMatrix(testing$classe, predgbm)

predrf = predict(moderf, testing)
confrf <- confusionMatrix(testing$classe, predrf)

```

```{r showmatrix, echo=FALSE}
confgbm
confrf
```

In this comparison, the overall accuracy of gbm model is `r round(confgbm$overall["Accuracy"]*100, 2)`%, and the overall accuracy of random forest model is `r round(confrf$overall["Accuracy"]*100, 2)`%. So it's obvious that the rf model is more accurate in this study.


### So finally, the rf prediction model is used to predict 20 different test cases.
```{r}
if(!file.exists("../pml-testing.csv")) {
        download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",
                      destfile = "../pml-testing.csv")
}

validation <- read.csv("../pml-testing.csv" ,header = TRUE,  na.strings = c("","NA","#DIV/0!"))

colnumber <- dim(intraining)[2]
validation <- subset(validation, select = names(intraining[,-colnumber]))

results <- predict(moderf, validation)
results <- paste(c(1:20), results, sep = ".")
```

## Results
```{r showresult, echo = FALSE}
results
```
